An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-30 VLDB2026/main.tex
1

\textit{\textbf{Index construction time.}}
\textcolor{orange}{Among the regular-scale datasets shown in the left part of Figure~\ref{fig:build_time_comparison_query1}, the methods with longer build times are mainly PASE, VBASE, and ACORN-$\gamma$. PASE and VBASE only support single-threaded builds. PASE is the slowest because its on-demand loading and caching strategies frequently trigger disk I/O and page scheduling, which seriously affects efficiency. VBASE uses full-memory builds and is relatively faster. ACORN-$\gamma$  has a large computational overhead because it needs to expand the neighbor list and evaluate a large number of candidates. The build time on some datasets even exceeds that of single-threaded VBASE. }

\textcolor{orange}{On large-scale dataset Deep100M, besides the three methods already mentioned, we also observe that algorithms based on Vamana (such as FilteredDiskANN and UNG) exhibit significantly increased construction time. Vamana adopts a two-pass graph construction and optimization strategy (greedy search + robust pruning), resulting in extremely high computational overhead. 
In contrast, the construction time of IVF-based algorithms grows much more slowly with dataset size, thanks to their shard-based design and high parallel efficiency, making them more scalable and advantageous in large-scale scenarios.}

@@

\textit{\textbf{ \textcolor{orange}{Search peak memory.}}}
%  \textcolor{orange}{As shown in Figure~\ref{fig:search_memory_mb_comparison}, we can find that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the peak memory usage of database is comparatively lower.}
%\textcolor{orange}{For most algorithms, the query phase requires loading both the index and the dataset into memory, so its peak memory usage is usually comparable to that of the construction phase. However, as shown in the Figure~\ref{fig:search_memory_mb_comparison}, CAPS exhibits significantly higher peak memory during querying due to the need to dynamically maintain large candidate sets. In contrast, Milvus has lower peak memory usage because it does not fully load the index and incurs additional metadata overhead during construction.
\textcolor{orange}{
Most algorithms load both the index and the dataset into memory during the build and query phases, and the construction process typically generates a large amount of temporary data. As a result, peak memory usage during querying is generally slightly lower than during construction. However, exceptions exist. As shown in Figure~\ref{fig:search_memory_mb_comparison}, CAPS incurs significantly higher peak memory usage during querying due to the need to dynamically maintain a large candidate set.}

@@

		\textcolor{orange}{
	In contrast, WinFilter relies on a tree-based structure, where a separate nearest neighbor graph must be built for each node. This design introduces redundant computation and high construction cost. However, on large-scale datasets, the preprocessing phase of WinFilter is replaced in the experiments by the base algorithm Vamana-WST, thus avoiding significant time overhead.}

>>>>

\textit{\textbf{Index construction time.}}
\textcolor{orange}{Among the regular-scale datasets shown in the left part of Figure~\ref{fig:build_time_comparison_query1}, the methods with longest build times are mainly PASE, VBASE, and ACORN-$\gamma$. PASE and VBASE only support single-threaded builds. PASE is the slowest because its on-demand loading and caching strategies frequently trigger disk I/O and page scheduling, which seriously affects efficiency. VBASE uses full-memory builds and is relatively faster. ACORN-$\gamma$  has a high computational overhead because it needs to expand the neighbor list and evaluate a large number of candidates. Its build time on some datasets even exceeds that of single-threaded VBASE. }

\textcolor{orange}{On the large-scale dataset Deep100M, besides the three methods already mentioned, we also observe that algorithms based on Vamana (such as FilteredDiskANN and UNG) exhibit significantly increased construction time. Vamana adopts a two-pass graph construction and optimization strategy (greedy search + robust pruning), resulting in extremely high computational overhead. In contrast, the construction time of IVF-based algorithms grows much more slowly with dataset size, thanks to their shard-based design and high parallel efficiency, making them more scalable and advantageous in large-scale scenarios.}

@@

\textit{\textbf{Search peak memory.}}
%  \textcolor{orange}{As shown in Figure~\ref{fig:search_memory_mb_comparison}, we can find that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the peak memory usage of database is comparatively lower.}
%\textcolor{orange}{For most algorithms, the query phase requires loading both the index and the dataset into memory, so its peak memory usage is usually comparable to that of the construction phase. However, as shown in the Figure~\ref{fig:search_memory_mb_comparison}, CAPS exhibits significantly higher peak memory during querying due to the need to dynamically maintain large candidate sets. In contrast, Milvus has lower peak memory usage because it does not fully load the index and incurs additional metadata overhead during construction.
\textcolor{orange}{
Most algorithms load both the index and the dataset into memory during the build and query phases, and the construction process typically generates a large amount of temporary data. As a result, peak memory usage during querying is generally slightly lower than during construction. However, exceptions exist. As shown in Figure~\ref{fig:search_memory_mb_comparison}, CAPS incurs significantly higher peak memory usage during querying due to the need to dynamically maintain a large candidate set.}

@@

		\textcolor{orange}{
	In contrast, WinFilter relies on a tree-based structure, where a separate nearest neighbor graph must be built for each node. This design introduces redundant computation and high construction cost. However, on large-scale datasets, the preprocessing phase of WinFilter was replaced with a base algorithm in the experiments, thus avoiding significant time overhead.}
